---
title: "Survey Data Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## Load Libraries
```{r}
library(tidyverse)
library(readxl)
library(DT)
library(dplyr)
```

## Load Data
```{r}
file_path <- "SurveyData.xlsx"
survey_main <- read_excel(file_path, sheet = 1)
family_members <- read_excel(file_path, sheet = "family_members")
earning_members <- read_excel(file_path, sheet = "earning_members")
family_responsibilities <- read_excel(file_path, sheet = "family_responsibilities")
health_details <- read_excel(file_path, sheet = "health_details")
```

## Data Overview
### Main Survey Data
```{r}
dim(survey_main)
glimpse(survey_main)
DT::datatable(head(survey_main))
```

### Family Members
```{r}
dim(family_members)
glimpse(family_members)
DT::datatable(head(family_members))
```

### Earning Members
```{r}
dim(earning_members)
glimpse(earning_members)
DT::datatable(head(earning_members))
```

### Family Responsibilities
```{r}
dim(family_responsibilities)
glimpse(family_responsibilities)
DT::datatable(head(family_responsibilities))
```

### Health Details
```{r}
dim(health_details)
glimpse(health_details)
DT::datatable(head(health_details))
```

## Summary Statistics
### Age Distribution
```{r}
survey_main %>%
  ggplot(aes(x = age)) + 
  geom_histogram(bins = 20, fill = "blue", alpha = 0.7) +
  labs(title = "Age Distribution of Farmers", x = "Age", y = "Count")
```

### Gender Distribution
```{r}
survey_main %>%
  count(gender) %>%
  ggplot(aes(x = gender, y = n, fill = gender)) + 
  geom_bar(stat = "identity", alpha = 0.7) +
  labs(title = "Gender Distribution", x = "Gender", y = "Count")
```

### Income Analysis
```{r}
earning_members %>%
  ggplot(aes(x = monthly_income)) + 
  geom_histogram(bins = 15, fill = "red", alpha = 0.7) +
  labs(title = "Income Distribution", x = "Monthly Income", y = "Count")
```

### Family Size Distribution
```{r}
family_members %>%
  #count(farmer_id) %>%
  ggplot(aes(x = n)) +
  geom_histogram(bins = 10, fill = "purple", alpha = 0.7) +
  labs(title = "Family Size Distribution", x = "Number of Family Members", y = "Count")
```

### Health Issues Analysis
```{r}
health_details %>%
  count(health_issue) %>%
  ggplot(aes(x = reorder(health_issue, -n), y = n, fill = health_issue)) +
  geom_bar(stat = "identity", alpha = 0.7) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Health Issues Reported", x = "Health Issue", y = "Count")
```

## Conclusion
- The dataset provides insights into the demographic, economic, and health status of affected families.
- Individual variable analysis provides deeper understanding of each factor.
- Further analysis can explore correlations between variables and socioeconomic impacts.